Bluetooth Loader for mruby Bytecode \\ in Multi-VM Environment

### 2016-04-06 EMSOFT2016/diff/0406_paper.tex
update

\title{mruby Bytecode Loader Using Bluetooth in Multi-VM Environment}
\numberofauthors{3}
\author{
\alignauthor
Takuro Yamamoto\\
    \affaddr{Graduate School of Engineering Science, Osaka University}\\
    \affaddr{Osaka, JAPAN}\\
       % \affaddr{Wallamaloo, New Zealand}\\
       % \email{trovato@corporation.com}
\alignauthor
Hiroshi Oyama\\
    \affaddr{OKUMA Corporation}\\
    \affaddr{Aichi, JAPAN}
       % \affaddr{Dublin, Ohio 43017-6221}\\
       % \email{webmaster@marysville-ohio.com}
\alignauthor
Takuya Azumi\\
    \affaddr{Graduate School of Engineering Science, Osaka University}\\
    \affaddr{Osaka, JAPAN}\\
       % \affaddr{Hekla, Iceland}\\
       % \email{larst@affiliation.org}
}
%\author{Takuro Yamamoto}
%\date{\today}

\begin{document}
\maketitle
\begin{abstract}
In recent years, the productivity of embedded systems has become a problem due to their complexity and large-scale.
For the purpose of improving the productivity for embedded software development, the mruby on TECS framework has been proposed that is applied mruby (Lightweight Ruby) and supports component-based development.
In the current mruby on TECS, the mruby programs have to be compiled and linked every time the programs are modified because the mruby bytecodes are incorporated in the platform.
Moreover, while the framework supports multi-VM, developers need to be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs in concurrent or/and parallel.
To improve the development efficiency, this paper proposes an mruby bytecode loader using Bluetooth as an extension of mruby on TECS.
The loader executes two mruby bytecodes, mruby application bytecode and mruby library bytecode.
mruby application bytecode modified frequently is sent from a host to a target device by developers.
mruby library bytecode modified infrequently is preserved beforehand with the platform in a storage/ROM device at the time of the first compilation.
In addition, multiple mruby programs cooperatively run in the proposed framework.
A RiteVM scheduler makes multitasking processing more easy-to-use than that of mruby on TECS.
Synchronization of starting multiple tasks is also implemented with Eventflag. 
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}

\keywords{embedded software; script language; component based development;}
\section{Introduction}
In these days, embedded systems have been required the high-quality and the high-performance.
Due to this trend, the complexity of embedded systems also increases and the scale is larger.
For example, IoT (Internet of Things) applications.
The low production cost and the short developing period of time are also needed.

An approach of efficient software development is to use component-based techniques.
CBD (Component-Based Development) is a design technique for constituting reusable components \cite{par:Crnkovic}.
Complex and large scale software systems can be developed efficiently using component-based techniques.
That is because software componentization provides high-reusability.
Verification of component-based systems has been researched \cite{par:Blaming} \cite{par:Verification}.
It also makes a system flexible in extensions and specification changes.
There are TECS \cite{par:TECS}, AUTOSAR \cite{url:AUTOSAR}, and SaveCCM \cite{par:SAVEapproach} as a typical component-based development for embedded systems

In addition, another approach of efficient software development is to develop with script languages.
Most of current software are programmed in C language, and the development with C takes a high cost and more time to develop.
Script languages make software engineering more efficient and shorten a development period because script languages have high-productivity from their usability.
Java script, Perl, Python, Lua and Ruby are well-known as representative script languages.

Although script languages are easy to use and read, their execution time are slower than C language's.
For embedded systems, the real-time properties such as estimation of worst-execution time are very important factors.
Therefore, it is difficult to apply the script languages to embedded systems.

mruby on TECS is a component-based framework for running script program \cite{par:mrubyonTECS}.
It is integrated two technologies.
One is mruby, which is a script language for embedded systems \cite{par:mruby}, \cite{url:mruby}.
The other is TECS (TOPPERS Embedded Component System), which is a component-based framework for embedded systems \cite{par:TECS} \cite{url:TOPPERS}.
mruby on TECS supports to effectively run mruby script language on embedded systems.
mruby on TECS also makes execution time 100 times faster than that of mruby.

%This paper proposes two additional features of mruby on TECS, mruby bytecode loader using Bluetooth and RiteVM scheduler.
mruby on TECS has several problems at present.
One of the problems is low software development efficiency.
mruby on TECS only supports a storage/ROM device to load mruby programs. 
%in the platform for LEGO MINDSTORMS EV3 \cite{par:EV3}.
For example, an SD card should be inserted and pulled out repeatedly, or a ROM should be rewritten if mruby programs are modified.
Developers should also restart an OS on a target device.
%Moreover, in the current system if developers attempt to run multiple tasks with the same priority, a task with first execution only runs and all other tasks do not run unless developers call the OS's function.
%In addition developers must prepare as many VMs as the number of tasks because one RiteVM supports only one task.
Moreover, although mruby on TECS has supported the multi-VM, developers need to call OS's function in order to execute multiple tasks.
These problem are a big burden on developers.

This paper proposes an extended framework of mruby on TECS: an mruby bytecode loader using Bluetooth and a RiteVM scheduler for fairly executing mruby programs.
To improve the development efficiency, the mruby bytecode loader using Bluetooth enables developers to write the platform in a storage device only once at the beginning and to transfer mruby application codes from a host to a target device.
The RiteVM scheduler enables developers to run multitasking programs more easily than mruby on TECS.
%Specifically, a periodic handler of switching tasks is implemented.
%This paper evaluates the proposed framework in term of data transfer rate, overhead of periodic handler and execution time.

{\mybf Contribution}: The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item {\mybf To improve the software development efficiency:} Developers do not need to rewrite a storage/ROM device and also to restart an OS.
The loader supports the continuous loading, which saves the Bluetooth set-up time.
Therefore, the mruby bytecode loader using Bluetooth helps developers develop software.
\item {\mybf To effectively execute multiple mruby programs in concurrent or/and parallel:} Developers can execute multiple tasks without the knowledge of an RTOS because the RiteVM scheduler switches tasks cyclically.
\item {\mybf To synchronize the execution of multiple RiteVM tasks:} The proposed framework provides the synchronization of multiple RiteVM tasks (mruby applications).
\item {\mybf To focus on the benefits of component-based developments:} This paper shows the specific examples for the benefits of component-based development. 
\end{enumerate}

@@

\section{Background}
\label{sec:Background}
Figure \ref{fig:proposed} shows the system model of the proposed framework.
The bytecodes are transferred from the host to the target device with Bluetooth.
The RiteVMs and mruby library are assumed to be prepared in advance.
Each bytecode is allocated to a RiteVM, respectively.
Developers can run bytecodes transferred with Bluetooth in multitask.

@@

mruby is suitable for embedded systems because of execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
A RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by a RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.eps}
    \vspace{1mm}
    \caption{Mechanism of mruby/RiteVM}
    \vspace{1mm}
    \label{fig:mruby}
\end{figure}

\subsection{TECS}
\label{sec:TECS}
TECS (TOPPERS Embedded Component System) is a component system suitable for embedded systems.
TECS helps increase productivity and reduce development duplication because the reusability of software is improved.
TECS also helps decrease complexity of the system because the generated component diagram can visualize the structure of whole software.

The component deployment and composition in TECS are statically performed, which gives optimization.
As a result, the overhead of execution time and memory consumption can be reduced.
There are other features of TECS, implementation in C language, source level portability, fine-grained component, etc.

@@

\subsubsection{Component Description}
The description of a component in TECS is divided into three parts, {\myit signature}, {\myit celltype}, and build description.
TECS code is written in .cdl (component description language) file.
The component description is mentioned with an example shown in Figure \ref{fig:component} as follow.

\begin{description}
    \item[{\mybf Signature Description}]\mbox{}\\
        The {\myit signature} description defines an interface of a {\myit cell}.
        A {\myit signature} name is described following the keyword {\myit signature}.
        It also has the prefix ``s".
        In this way, a {\myit signature} is defined such as sMotor shown in Figure \ref{signature}.
        To make the definition of an interface clear, specifiers such as in and out are used in TECS.
        [in] and [out] represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor{
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\vspace{1mm}
\caption{Signature Description}
\vspace{1mm}
\label{signature}
\end{figure}
    \item[{\mybf Celltype Description}]\mbox{}\\
        The {\myit celltype} description defines {\myit entry} ports, {\myit call} ports, attributes, and valuables of a {\myit celltype}.
        An example of a {\myit celltype} description is shown in Figure \ref{celltype}.
        A {\myit celltype} name following the keyword {\myit celltype} with the prefix ``t" and elements of a {\myit celltype} is described.
        To define {\myit entry} ports, a {\myit signature} such as sMotor, and an {\myit entry} port name such as eMotor follow the keyword {\myit entry}.
        In the same way, {\myit call} ports can be defined.
        Attributes and valuables follow the keyword {\myit attr} and {\myit var} respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller{
    call sMotor cMotor;
};
celltype tMotor{
    entry sMotor eMotor;
    attr{
        int32_t port;
    };
    var{
        int32_t currentSpeed = 0;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description}
\vspace{1mm}
\label{celltype}
\end{figure}
    \item[{\mybf Build Description}]\mbox{}\\
        The build description is used to instantiate {\myit cell}s and connect {\myit cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\myit celltype} name such as tMotor, and a {\myit cell} name such as Motor follow the keyword {\myit cell}.
        To compose {\myit cell}s, a {\myit call} port, a {\myit signature}, a {\myit entry} port in order are described.
        In this example, a {\myit entry} port eMotor in a {\myit cell} Motor is connected to a {\myit call} port cMotor in a {\myit cell} Caller.
\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor{
};
cell tCaller Caller{
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\vspace{1mm}
\caption{Build Description}
\vspace{1mm}
\label{build}
\end{figure}

@@

The use of Ruby on embedded devices provides the benefit of productivity and maintainability due to the ease to use and read.
On the other hands, it is necessary to implement parts of applications in C language in order to manipulate actuators and sensors, and also make a critical section of codes run quickly.

@@

\section{Design and Implementation}
\label{sec:Design and Implementation}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
Each mruby application bytecode transferred from the host to the target deviceis received by the loader in the RiteVM.
The RiteVM reads the transferred bytecode and executes it with libraries.
The mruby applications run at the same time because of synchronization processing.
The RiteVM scheduler switches RiteVM tasks as multiple tasks can run in concurrent.
This section explains these proposed functionalities.

@@

\subsection{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
\footnote{
mruby bytecode loader is intended to improve the development efficiency, therefore software developers should use it for development phase.
Complete software should be compiled and linked in the storage/ROM device beforehand.
}
This section describes an additional functionality of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, the platform including mruby bytecodes are saved in a storage/ROM device.
Developers must rewrite the storage/ROM device every time the application programs are modified.
The OS on the target device should be also restarted.
It causes a low development efficiency to repeat that.
An mruby bytecode loader using Bluetooth makes the developers' burden decrease.
%The work that an SD card is in and out needs to be done just once in the beginning.
Developers should just once carry out the work to connect the storage/ROM device and start the OS.

@@

%The development flow in an mruby bytecode loader using Bluetooth is shown in Figure \ref{fig:bluetooth_loader}.
In the proposed framework, the platform including RiteVMs and mruby library is compiled and copied in the storage/ROM device at the first.
%These are assumed not to be modified.
%The binary data transferred with Bluetooth is the bytecode of the main source.
In the host, the mruby application programs (.rb) are edited and compiled into the bytecodes (.mrb) by an mruby compiler.
The generated bytecodes are transferred from the host to the target device with Bluetooth.

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=8cm,clip]{figure/bluetooth_loader.eps}
%     \vspace{1mm}
% \caption{Development Flow in mruby bytecode loader using Bluetooth}
%     \vspace{1mm}
% \label{fig:bluetooth_loader}
% \end{figure}

\subsubsection{Component of RiteVM with mruby bytecode loader using Bluetooth}
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.eps}
    \vspace{1mm}
\caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \vspace{1mm}
\label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with mruby bytecode loader using Bluetooth as TECS component.
The component is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages a RiteVM configuration such as automatically generation the mruby library bytecode.
This generated bytecode is prepared beforehand in the storage/ROM device, and different from a bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and RiteVMBluetooth1 {\myit cell}s.
The MrubyTask1 {\myit cell} is a componentized task of the RTOS (TOPPERS/HRP2).
TOPPERS/HRP2 is described in \cite{url:HRP2}, \cite{par:hr-tecs}.
The RiteVMBluetooth1 is the component of RiteVM with mruby bytecode loader using Bluetooth.
A bytecode in the host is transferred and received at the top of the component.
In this framework, ZMODEM \cite{par:zmodem} is used as a binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.eps}
    \vspace{1mm}
\caption{Process Flow of mruby bytecode loader using Bluetooth}
    \vspace{1mm}
\label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth{
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr{
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var{
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description for RiteVM with mruby bytecode loader using Bluetooth}
\vspace{1mm}
\label{celltype_mrubybluetooth}
\end{figure}

@@

\end{lstlisting}
\vspace{1mm}
\caption{Main code for RiteVM with mruby bytecode loader using Bluetooth}
\vspace{1mm}
\label{maincode_mrubybluetooth}
\end{figure}

Figure \ref{fig:control_flow} shows the process flow of executing mruby program in the component of RiteVM with mruby bytecode loader using Bluetooth such as RiteVMBluetooth1.
The concrete main code of tRiteVMBluetooth is shown in Figure \ref{maincode_mrubybluetooth}.

First, the mruby bytecode loader receives the mruby application bytecode transferred from the host (Figure \ref{fig:control_flow}:A, Lines 6-7 in Figure \ref{maincode_mrubybluetooth}).
The bytecode is stored in the component variable such as VAR\_irepApp shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

@@

Finally, the mruby task runs (Figure \ref{fig:control_flow}:E, Figure \ref{maincode_mrubybluetooth}:20-22).
When the mruby application is modified, only the bytecode of the modified application should be transferred.
mruby libraries need not be touched because libraries are not normally changed.

@@

The RiteVM scheduler is a fair scheduler and implemented as a TECS component in the proposed framework.
Therefore, when developers create software in priority-based scheduling, the RiteVM scheduler can be easily removed.
\subsubsection{Design of RiteVM Scheduler}
A RiteVM scheduler is a periodic handler, and calls {\myit rotateReadyQueue}, a service call of $\mu$ITRON to switch tasks with the same priority.
In other words, the RiteVM scheduler calls {\myit rotateReadyQueue} cyclically.
The design of the RiteVM scheduler is shown in Figure \ref{fig:rotateReadyQueue}. 
{\myit rotateReadyQueue} is described as follows.

@@

The {\myit rotateReadyQueue} can be performed if the number of tasks is more than two.
For example, three tasks are in the order: task 1, 2, and 3.
In this case, the order is rotated, task 2, 3 and, 1, when the {\myit rotateReadyQueue} is called.

>>>>

\title{Bluetooth Loader for mruby Bytecode \\ in Multi-VM Environment}
\numberofauthors{3}
\author{
\alignauthor
Takuro Yamamoto\\
    \affaddr{Graduate School of Engineering Science, Osaka University}\\
    \affaddr{Osaka, JAPAN}\\
       % \affaddr{Wallamaloo, New Zealand}\\
       % \email{trovato@corporation.com}
\alignauthor
Hiroshi Oyama\\
    \affaddr{OKUMA Corporation}\\
    \affaddr{Aichi, JAPAN}\\
       % \affaddr{Dublin, Ohio 43017-6221}\\
       % \email{webmaster@marysville-ohio.com}
\alignauthor
Takuya Azumi\\
    \affaddr{Graduate School of Engineering Science, Osaka University}\\
    \affaddr{Osaka, JAPAN}\\
       % \affaddr{Hekla, Iceland}\\
       % \email{larst@affiliation.org}
}
%\author{Takuro Yamamoto}
%\date{\today}

\begin{document}
\maketitle
\begin{abstract}
In recent years, the productivity of embedded systems has a problem due to their complexity and large-scale.
For the purpose of improving the productivity for embedded software development, the mruby on TECS framework has been proposed that is applied mruby (Lightweight Ruby) and supports component-based development.
In the current mruby on TECS, the mruby programs have to be compiled and linked every time the programs are modified because the mruby bytecodes are incorporated in the platform.
Moreover, while the framework supports multi-VM, developers need to be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs in concurrent or/and parallel.
To improve the development efficiency, this paper proposes a Bluetooth loader for mruby bytecode as an extension of mruby on TECS.
The loader executes two mruby bytecodes, mruby application bytecode and mruby library bytecode.
mruby application bytecode is modified frequently and sent from a host to a target device by developers.
mruby library bytecode modified infrequently is preserved beforehand with the platform in a storage/ROM device at the time of the first compilation.
In addition, multiple mruby programs cooperatively run in the proposed framework.
A RiteVM scheduler makes multitasking processing more easy-to-use than that of mruby on TECS.
Synchronization of starting multiple tasks is also implemented with Eventflag. 
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}

\keywords{embedded software; script language; component-based development;}
\section{Introduction}
In these days, embedded systems have been required the high-quality and the high-performance.
Due to this trend, the complexity of embedded systems also increases and the scale is larger.
For example, IoT (Internet of Things) applications.
The low production cost and the short developing period of time are also needed.

An approach of efficient software development is to use component-based techniques.
CBD (Component-Based Development) is a design technique for constituting reusable components \cite{par:Crnkovic}.
Complex and large scale software systems can be developed efficiently using component-based techniques.
That is because software componentization provides high-reusability.
Verification of component-based systems has been researched \cite{par:Blaming} \cite{par:Verification}.
It also visualize the entire system due to component diagrams, and makes a system flexible in extensions and specification changes.
There are TECS \cite{par:TECS}, AUTOSAR \cite{url:AUTOSAR}, and SaveCCM \cite{par:SAVEapproach} as a typical component-based development for embedded systems

In addition, another approach of efficient software development is to develop with script languages.
Most of current software are programmed in C language.
Development with C language makes the code size increase, and takes a high cost and more time to develop.
Script languages make software engineering more efficient and shorten a development period because script languages have high-productivity from their usability.
Ruby, JavaScript, Perl, Python, and Lua are well-known as representative script languages.

Although script languages are easy to use and read, their execution time are larger than C language's.
For embedded systems, the real-time properties such as estimation of worst-case execution time are very important factors.
Therefore, it is difficult to apply the script languages to embedded systems.

To improve the above probles, mruby on TECS has been proposed, which is a component-based framework for running script program \cite{par:mrubyonTECS}.
It is integrated two technologies.
One is mruby, which is a script language for embedded systems \cite{par:mruby}, \cite{url:mruby}.
The other is TECS (TOPPERS Embedded Component System), which is a component-based framework for embedded systems \cite{par:TECS} \cite{url:TOPPERS}.
mruby on TECS supports to effectively run mruby script language on embedded systems.
mruby on TECS also makes execution time 100 times faster than that of mruby.

%This paper proposes two additional features of mruby on TECS, Bluetooth loader for mruby bytecode and RiteVM scheduler.
mruby on TECS has several problems at present.
One of the problems is low software development efficiency.
mruby on TECS only supports a storage/ROM device to load mruby programs. 
%in the platform for LEGO MINDSTORMS EV3 \cite{par:EV3}.
For example, an SD card should be inserted and pulled out repeatedly, or a ROM should be rewritten if mruby programs are modified.
Developers should also restart an RTOS on a target device.
%Moreover, in the current system if developers attempt to run multiple tasks with the same priority, a task with first execution only runs and all other tasks do not run unless developers call the OS's function.
%In addition developers must prepare as many VMs as the number of tasks because one RiteVM supports only one task.
Moreover, although mruby on TECS has supported the multi-VM, developers need to call OS's function in order to execute multiple tasks.
These problems makes the burden heavy for developers.

This paper proposes an extended framework of mruby on TECS: a Bluetooth loader for mruby bytecode and a RiteVM scheduler for fairly executing mruby programs.
To improve the development efficiency, the proposed framework enables developers to implement the platform in a storage device only once at the beginning and to transfer mruby application programs from a host to a target device using the Bluetooth Loader.
A RiteVM is a virtual machine for mruby.
The RiteVM scheduler manages the execution of multiple RiteVMs, and enables developers to program multitasking more easily than mruby on TECS.
%Specifically, a periodic handler of switching tasks is implemented.
%This paper evaluates the proposed framework in term of data transfer rate, overhead of periodic handler and execution time.

{\mybf Contribution}: The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item {\mybf To improve the software development efficiency:} Developers do not need to rewrite a storage/ROM device and also to restart an RTOS.
The loader supports the continuous loading, which saves the Bluetooth set-up time.
Therefore, the Bluetooth loader for mruby bytecode helps developers develop software.
\item {\mybf To effectively execute multiple mruby programs in concurrent or/and parallel:} Developers can implement multiple tasks without the knowledge of an RTOS because the RiteVM scheduler switches tasks cyclically.
\item {\mybf To synchronize the execution of multiple RiteVM tasks:} The proposed framework provides the synchronization of multiple RiteVM tasks (mruby applications).
\item {\mybf To focus on the benefits of component-based developments:} This paper shows the specific examples for the benefits of component-based development. 
\end{enumerate}

@@

\section{Background}
\label{sec:Background}
Figure \ref{fig:proposed} shows the system model of the proposed framework.
The bytecodes are transferred from the host to the target device with Bluetooth.
The RiteVMs and mruby library are assumed to be prepared in advance.
Each bytecode is allocated to a RiteVM, respectively.
Bytecodes transferred with Bluetooth from the host can run in multitask.

@@

mruby is suitable for embedded systems because of execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
A RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by a RiteVM.
The bytecodes can run on a RiteVM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.eps}
    \vspace{1mm}
    \caption{Mechanism of mruby/RiteVM}
    \vspace{1mm}
    \label{fig:mruby}
\end{figure}

\subsection{TECS}
\label{sec:TECS}
TECS is a component system suitable for embedded systems.
TECS helps increase productivity and reduce development duplication because the reusability of software is improved.
TECS also helps developers understand the system because the generated component diagram can visualize the structure of whole software.

The component deployment and composition in TECS are statically performed, which gives optimization.
As a result, the overhead of connection of components and the memory consumption can be reduced.
There are other features of TECS, implementation in C language, source level portability, fine-grained component, etc.

@@

\subsubsection{Component Description}
The description of a component in TECS is divided into three parts, {\myit signature}, {\myit celltype}, and build description.
TECS code is written in CDL (component description language) file.
The component description is mentioned with an example shown in Figure \ref{fig:component} as follow.

\begin{description}
    \item[{\mybf Signature Description}]\mbox{}\\
        The {\myit signature} description defines an interface of a {\myit cell}.
        A {\myit signature} name is described following the keyword {\myit signature}.
        It also has the prefix ``s".
        In this way, a {\myit signature} is defined such as sMotor shown in Figure \ref{signature}.
        To make the definition of an interface clear, specifiers such as in and out are used in TECS.
        [in] and [out] represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor {
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\vspace{1mm}
\caption{Signature Description}
\vspace{1mm}
\label{signature}
\end{figure}
    \item[{\mybf Celltype Description}]\mbox{}\\
        The {\myit celltype} description defines {\myit entry} ports, {\myit call} ports, attributes, and variable of a {\myit celltype}.
        An example of a {\myit celltype} description is shown in Figure \ref{celltype}.
        A {\myit celltype} name following the keyword {\myit celltype} with the prefix ``t" and elements of a {\myit celltype} is described.
        To define {\myit entry} ports, a {\myit signature} such as sMotor, and an {\myit entry} port name such as eMotor follow the keyword {\myit entry}.
        In the same way, {\myit call} ports can be defined.
        Attributes and variables follow the keyword {\myit attr} and {\myit var} respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller {
    call sMotor cMotor;
};
celltype tMotor {
    entry sMotor eMotor;
    attr {
        int32_t port;
    };
    var {
        int32_t currentSpeed = 0;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description}
\vspace{1mm}
\label{celltype}
\end{figure}
    \item[{\mybf Build Description}]\mbox{}\\
        The build description is used to instantiate {\myit cell}s and connect {\myit cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\myit celltype} name such as tMotor, and a {\myit cell} name such as Motor follow the keyword {\myit cell}.
        To compose {\myit cell}s, a {\myit call} port, a {\myit signature}, a {\myit entry} port in order are described.
        In this example, a {\myit entry} port eMotor in a {\myit cell} Motor is connected to a {\myit call} port cMotor in a {\myit cell} Caller.
        The {\myit C\_EXP} is the keyword to use macros defined in C files.

\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor {
    port = C_EXP("PORT_A");
};
cell tCaller Caller {
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\vspace{1mm}
\caption{Build Description}
\vspace{1mm}
\label{build}
\end{figure}

@@

The use of Ruby on embedded devices provides the benefit of productivity and maintainability due to the ease to use and read.
On the other hand, it is necessary to implement parts of applications in C language in order to manipulate actuators and sensors, and also make a critical section of codes run quickly.

@@

\section{Design and Implementation}
\label{sec:Design and Implementation}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
Each mruby application bytecode transferred from the host to the target devices received by the loader in the RiteVM.
The RiteVM reads the transferred bytecode and executes it with libraries.
The mruby applications run at the same time because of synchronization processing.
The RiteVM scheduler switches RiteVM tasks as multiple tasks can run in concurrent.
This section explains these proposed functionalities.

@@

\subsection{Bluetooth Loader for mruby Bytecode}
\label{sec:Bluetooth loader for mruby bytecode}
This section describes an additional functionality of mruby on TECS, Bluetooth loader for mruby bytecode.
\footnote{
Bluetooth loader is intended to improve the development efficiency, therefore software developers should use it for development phase.
Complete software should be compiled and linked in the storage/ROM device beforehand.
}
In the current system, the platform including mruby bytecodes are saved in a storage/ROM device.
Developers must rewrite the storage/ROM device every time the application programs are modified.
The RTOS on the target device should be also restarted.
It causes a low development efficiency to repeat that.
A Bluetooth loader for mruby bytecode makes the developers' burden decrease.
%The work that an SD card is in and out needs to be done just once in the beginning.
Developers should just once carry out the work to connect the storage/ROM device and start the RTOS.

@@

%The development flow in a Bluetooth loader for mruby bytecode is shown in Figure \ref{fig:bluetooth_loader}.
In the proposed framework, the platform including RiteVMs and mruby library is compiled and copied in the storage/ROM device at the first.
%These are assumed not to be modified.
%The binary data transferred with Bluetooth is the bytecode of the main source.
In the host, the mruby application programs (.rb) are edited and compiled into the bytecodes (.mrb) by an mruby compiler.
The generated bytecodes are transferred from the host to the target device with Bluetooth.

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=8cm,clip]{figure/bluetooth_loader.eps}
%     \vspace{1mm}
% \caption{Development Flow in Bluetooth loader for mruby bytecode}
%     \vspace{1mm}
% \label{fig:bluetooth_loader}
% \end{figure}

\subsubsection{Component of RiteVM with Bluetooth loader for mruby bytecode}
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.eps}
    \vspace{1mm}
\caption{Component Diagram of Bluetooth loader for mruby bytecode}
    \vspace{1mm}
\label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with Bluetooth loader for mruby bytecode as TECS component.
The component is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages a RiteVM configuration such as automatically generating the mruby library bytecode.
This generated bytecode is prepared beforehand in the storage/ROM device, and different from a bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and RiteVMBluetooth1 {\myit cell}s.
The MrubyTask1 {\myit cell} is a componentized task of the RTOS (TOPPERS/HRP2).
TOPPERS/HRP2 is described in \cite{url:HRP2}, \cite{par:hr-tecs}.
The RiteVMBluetooth1 is the component of RiteVM with Bluetooth loader for mruby bytecode.
A bytecode in the host is transferred and received at the top of the component.
In this framework, ZMODEM \cite{par:zmodem} is used as a binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.eps}
    \vspace{1mm}
\caption{Process Flow of Bluetooth loader for mruby bytecode}
    \vspace{1mm}
\label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth1 {
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr {
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var {
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description for RiteVM with Bluetooth loader for mruby bytecode}
\vspace{1mm}
\label{celltype_mrubybluetooth}
\end{figure}

@@

\end{lstlisting}
\vspace{1mm}
\caption{Main code for RiteVM with Bluetooth loader for mruby bytecode}
\vspace{1mm}
\label{maincode_mrubybluetooth}
\end{figure}

Figure \ref{fig:control_flow} shows the process flow of executing mruby program in the component of RiteVM with Bluetooth loader for mruby bytecode such as RiteVMBluetooth1.
The concrete main code of tRiteVMBluetooth is shown in Figure \ref{maincode_mrubybluetooth}.

First, the Bluetooth loader receives the mruby application bytecode transferred from the host (Figure \ref{fig:control_flow}:A, Lines 6-7 in Figure \ref{maincode_mrubybluetooth}).
The bytecode is stored in the component variable such as VAR\_irepApp shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

@@

Finally, the mruby task runs (Figure \ref{fig:control_flow}:E, Lines 20-22 in Figure \ref{maincode_mrubybluetooth}).
When the mruby application is modified, only the bytecode of the modified application should be transferred.
mruby libraries need not be touched because libraries are not normally changed.

@@

The RiteVM scheduler is a fair scheduler and implemented as a TECS component in the proposed framework.
Therefore, when developers create software in priority-based scheduling, the RiteVM scheduler can be easily removed.

\subsubsection{Design of RiteVM Scheduler}
A RiteVM scheduler is a periodic handler and {\myit rotateReadyQueue}, a service call of $\mu$ITRON to switch tasks with the same priority, is implemented as the main process of the handler.
In other words, the RiteVM scheduler calls {\myit rotateReadyQueue} cyclically.
The design of the RiteVM scheduler is shown in Figure \ref{fig:rotateReadyQueue}. 
{\myit rotateReadyQueue} is described as follows.

@@

{\myit rotateReadyQueue} can be performed if the number of tasks is more than two.
For example, three tasks are in the order: task1, task2, and, task3.
In this case, the order is rotated, task2, task3, and, task1, when the {\myit rotateReadyQueue} is called.
